---
output: html_document
---
# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

*Show any code that is needed to*  

1. *Load the data (i.e. `read.csv()`)*  

2. *Process/transform the data (if necessary) into a format suitable for your analysis*

Reading the data:
```{r}
activityData <- read.csv('activity.csv')
completeActivityData <- activityData[complete.cases(activityData),]
dailyActivity <- split(completeActivityData, completeActivityData$date)
dailyActivity1 <- split(activityData, activityData$date)
intervalActivity <- split(activityData, activityData$interval)
```

## What is mean total number of steps taken per day?
*For this part of the assignment, youu can ignore the missing values in the dataset.*

1. *Make a histogram of the total number of steps taken each day*

2. *Calculate and report the **mean** and **median** total number of steps taken per day*


```{r}
  dailySteps <- unlist(lapply(dailyActivity, function(x){
          sum(x$steps)
        })
        )
  hist(dailySteps, main = "Histogram of the total numbers of daily steps")
```

For mean and median computation we use.
```{r} 
mean(dailySteps, na.rm=TRUE)
median(dailySteps, na.rm=TRUE)
```

## What is the average daily activity pattern?

```{r}
  intervalSteps <-numeric()
  trash <- lapply(intervalActivity, function(x){
                  intervalSteps <<- c(intervalSteps, mean(x$steps, na.rm=TRUE))
                  })
    plot(unique(activityData$interval), intervalSteps, type='l', xlab ="interval", ylab = "mean step number", main="timeseries plot of mean steps per interval")
```


## Imputing missing values

Handling missing values:
Calculation of rows with `NA` values:
```{r}
  rowsWithNa <- nrow(activityData) - nrow(completeActivityData)
  rowsWithNa
```

We reset the missing values with the mean of the respective 5 minute interval

```{r}
  newActivityData <- activityData
  naVec <- is.na(newActivityData)
  intervalIdx <- newActivityData[naVec, "interval"]
  newActivityData[naVec,"steps"] <- intervalSteps[intervalIdx/5 +1]
```


Histogram of steps taken per day:
```{r}
  newDailySteps <- numeric()
  newDailyActivity <- split(newActivityData, newActivityData$date)
  trash <- lapply(newDailyActivity, function(x){
        ds <-sum(x$steps)
        newDailySteps <<- c(newDailySteps, ds)
        })
  hist(newDailySteps, main = "Histogram of the total numbers of daily steps using the added values")
```

Let us calculate the daily mean and the median of the steps:
```{r} 
mean(newDailySteps, na.rm=TRUE)
median(newDailySteps, na.rm=TRUE)
```

## Are there differences in activity patterns between weekdays and weekends?

```{r}
  newActivityData$dayType <- as.factor(ifelse(weekdays( as.Date(newActivityData$date)) %in% c("Saturday","Sunday"), "Weekend", "Weekday"))
  #panel.xyplot(steps~interval|dayType, newActivityData)
  #densityplot(~steps|dayType, newActivityData)

```

